---
title: "Penguin Data EDA"
output: html_notebook
---

HW: Label the entire notebook with comments on what operation is each code chunk doing and what is the outcome

```{r}
library(tidyverse)
library(plotly)
Penguins<-read.csv("penguins_size.csv")
```

We see that the data has 344 rows and 7 columns

```{r}
Penguins
dim(Penguins)
```

Looking at summary stats
```{r}
summary(Penguins)
```
We notice that Species and island are read in as characters. We will convert them to factor.
```{r}
Penguins$species<-as.factor(Penguins$species)
Penguins$island<-as.factor(Penguins$island)
Penguins$sex<-as.factor(Penguins$sex)
```

Looking at summary again
```{r}
summary(Penguins)
```
We will remove the null values and also any erroneous values for sex  
```{r}
Clean_DF<-na.omit(Penguins)
Clean_DF<-Clean_DF%>%
  filter(sex == 'FEMALE' | sex == 'MALE')
```

This leaves us with the below summary stats:
```{r}
summary(Clean_DF)
```

```{r}
Clean_DF$sex<-as.factor(Clean_DF$sex)
```

```{r}
summary(Clean_DF)
```

```{r}
boxplot(Clean_DF[3:5])
```

```{r}
boxplot(Clean_DF$body_mass_g)
```

```{r}
p<-ggplot(Clean_DF, aes(sex, body_mass_g, fill=sex))+
  geom_boxplot()

ggplotly(p)
```

```{r}
p<-ggplot(Clean_DF, aes(sex, culmen_length_mm, fill=sex))+
  geom_boxplot()

ggplotly(p)
```


```{r}
p<-ggplot(Clean_DF, aes(sex, culmen_depth_mm, fill=sex))+
  geom_boxplot()

ggplotly(p)
```

```{r}
p<-ggplot(Clean_DF, aes(sex, flipper_length_mm, fill=sex))+
  geom_boxplot()

ggplotly(p)
```

Looking at body measures by Species

```{r}
p<-ggplot(Clean_DF, aes(species, body_mass_g, fill=species))+
  geom_boxplot()

ggplotly(p)
```

```{r}
p<-ggplot(Clean_DF, aes(species, culmen_length_mm, fill=species))+
  geom_boxplot()

ggplotly(p)
```


```{r}
p<-ggplot(Clean_DF, aes(species, culmen_depth_mm, fill=species))+
  geom_boxplot()

ggplotly(p)
```

```{r}
p<-ggplot(Clean_DF, aes(species, flipper_length_mm, fill=species))+
  geom_boxplot()

ggplotly(p)
```

```{r}
p<-ggplot(data = Clean_DF) + 
  geom_bar(mapping = aes(x = species, fill=species))

ggplotly(p)
```

```{r}
p<-ggplot(data = Clean_DF) + 
  geom_bar(mapping = aes(x = island, fill=species))

ggplotly(p)
```



```{r}
p<-ggplot(data = Clean_DF) + 
  geom_point(mapping = aes(x = culmen_length_mm, y = culmen_depth_mm,color = species, shape=island))

ggplotly(p)
```

```{r}
p<-ggplot(data = Clean_DF) + 
  geom_point(mapping = aes(x = culmen_length_mm, y = flipper_length_mm,color = species, shape=island))

ggplotly(p)
```

```{r}
p<-ggplot(data = Clean_DF) + 
  geom_point(mapping = aes(x = culmen_length_mm, y = body_mass_g, color = species, shape=island))

ggplotly(p)
```

```{r}
library(corrplot)
M<-cor(Clean_DF[3:6])
corrplot(M,method="color",addCoef.col = "white")
```

