---
title: "R Notebook"
output: html_notebook
---

Let's import the iris dataset and explore it

```{r}
data(iris)
```

we will first try to understand the structure of the data
```{r}
head(iris)
dim(iris)
```
We notice that the iris data has 5 columns and 150 observations/rows

```{r}
colnames(iris)
```
Let us look at some summary statistics
```{r}
summary(iris)
```
Another way of deriving these summary statistics is to do it one after the other using dedicated functions.
```{r}
min(iris$Sepal.Length)
quantile(iris$Sepal.Length,probs=0.25)
median(iris$Sepal.Length)
mean(iris$Sepal.Length)
quantile(iris$Sepal.Length,probs=0.75)
max(iris$Sepal.Length)
```
Here we notice that mode is not present in the summary. How do we compute mode? 

R does not give a predefined  mode function. hence we will write our own procedure 
```{r}
names(sort(-table(iris$Sepal.Length)))[1]
```
or function

```{r}
mode<-function(x){
  names(sort(-table(x)))[1]
}
```

```{r}
mode(iris$Sepal.Width)
```
we can also understand the spread of the data by looking at the standard deviation, variance, range or IQR
```{r}
sd(iris$Sepal.Length)
var(iris$Sepal.Length)
range(iris$Sepal.Length)[1]-range(iris$Sepal.Length)[2]
quantile(iris$Sepal.Length,probs=0.25)-quantile(iris$Sepal.Length,probs=0.75)
```


Now that we have understood what the distribution looks like, we could also try to look at the data in a visual manner

```{r}
hist(iris$Sepal.Length)
```

```{r}
hist(iris$Sepal.Width)
```

```{r}
hist(iris$Petal.Length)
```

```{r}
hist(iris$Petal.Width)
```
A boxplot shows the above descriptive stats visually and also shows outliers. An outlier is a point that is < = (Q1 - 1.5 * IQR) or > = (Q3 + 1.5 * IQR)

```{r}
boxplot(iris)
```

We can also get only the descriptives by using the .stats procedure
```{r}
boxplot.stats(iris$Sepal.Width)
```

```{r}
plot(iris$Species,iris$Sepal.Length)
```

```{r}
plot(iris$Species,iris$Sepal.Width)
```

```{r}
plot(iris$Species,iris$Petal.Length)
```

```{r}
plot(iris$Species,iris$Petal.Width)
```

Scatterplots and correlation

```{r}
plot(iris$Sepal.Length,iris$Petal.Length)
```

```{r}
plot(iris$Sepal.Width,iris$Petal.Width)
```

```{r}
setosa<-subset(iris,iris$Species=='setosa')
```

```{r}
plot(setosa$Sepal.Length,setosa$Petal.Length)
```

```{r}
plot(setosa$Sepal.Width,setosa$Petal.Width)
```

```{r}
plot(setosa$Sepal.Width,setosa$Sepal.Length)
```

Each scatter plot gives a sense of the correlation between two variables but doesn't give a value to it. Also computing each correlation pair for a large data set is a lot of work, hence we use the cor() function
```{r}
cor(setosa[1:4])
```

```{r}

```

