---
title: "6__ggplot"
output: html_notebook
---

Start by calling the tidyverse library. You will notice below that the tidyverse library has attached multiple packages. Tidyverse is a collection of packages developed by Rstudio that help doing advanced analytics in R.


```{r}
library(tidyverse)
data(iris)
```
We can make a scatter plot with colors based on a categorical variable
```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width,color  = Species))

```

We can fit a exponential smoothing curve for the scatter plot by the categorical variable
```{r}
ggplot(data = iris) + 
  geom_smooth(mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species, linetype= Species))
```

or even a linear model. More about linear models when we talk about regression
```{r}
ggplot(data = iris) + 
   geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width,color = Species))+
  geom_smooth(mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species), method = "lm")
```

or combine scatter plot and smoothing curves
```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width,color = Species))+
  geom_smooth(mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species, linetype= Species))

```

We can also better represent scatter plots by adding a little white noise to avoid overlap of points
```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width,color = Species), position = "jitter")

```

This is the same bar plot we saw earlier but with color.
```{r}
ggplot(data = iris) + 
  geom_bar(mapping = aes(x = Species, fill=Species))
```
However with ggplot2 you could use alternate aggregations for the y axis
```{r}
ggplot(iris) + 
  geom_bar(aes(Species, Petal.Length, colour = as.factor(Species),fill = as.factor(Species)), 
           position = "dodge", stat = "summary", fun.y = "mean")
```

or you could also plot summary stats
```{r}
ggplot(data = iris) + 
  stat_summary(
    mapping = aes(x = Species, y = Petal.Length, fill=Species),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

```{r}
ggplot(data = iris) + 
  stat_summary(
    mapping = aes(x = Species, y = Petal.Length),
    fun.data = "mean_sdl"
  )
```

