---
title: "6_2_dplyr"
output: html_document
---

Now that we have seen how to graphically analyse data. Let us also explore how we can use table manipulations to explore our data better
```{r}
library(tidyverse)
data(iris)
```

```{r}
iris
```

We can filter rows based on conditions
```{r}
#%>% --- pipe operator#
setosa<-iris%>% 
  filter(Species=="setosa")
```

```{r}
setosa
```


```{r}
not_setosa<-iris%>%
  filter(Species!="setosa")
print(not_setosa)
```

```{r}
A<-iris%>%
  filter(Petal.Length>5)
```

```{r}
print(A)
```



```{r}
also_not_setosa<-iris%>%
  filter(Species == "versicolor" | Species == "virginica")
```

We can select only the columns we need

```{r}
iris%>%
  select(Species, Petal.Length, Sepal.Length)
```

And we can also compute summaries for groups/categories
```{r}
iris%>%
  group_by(Species)%>%
  summarise(Total=sum(Petal.Length), count=n(), avg=mean(Petal.Length), median=median(Petal.Length))
```

or even derive new variables

```{r}
modified_iris<-iris%>%
  mutate(petal_sepal_ratio=Petal.Length/Sepal.Length)
```

```{r}
print(modified_iris)
```

or even derive new variables and further summarize them in a step by step manner

```{r}
summary(modified_iris)
```


```{r}
iris%>%
  mutate(petal_sepal_ratio=Petal.Length/Sepal.Length)%>%
  filter(petal_sepal_ratio>0.3 & petal_sepal_ratio<0.9)%>%
  group_by(Species)%>%
  summarise(avg_ratio=mean(petal_sepal_ratio))
```



