---
title: "Univariate Regression on Cars data"
output:
  html_document:
    df_print: paged
---

You can find a version of this document already knit and published on the cloud here: https://rpubs.com/sachinvarriar/cars_regression

#### alpha - intercept
#### beta(s) - coefficient(s)
#### X - independent variable/feature
#### Y - dependent variable/response

Y=f(X)
Y is dependent on the value of X. Think of substituting X in the fn f() to obtain Y

```{r}
data('cars')
summary(cars)
```
lm() stands for linear model. Why linear model? This performs a "linear" regression on the data and gives us the equation of the line that generalizes the relation between Y and X(s)

```{r}
model<-lm(speed~dist,cars)
```

```{r}
print(model)
```
Y=8.2839 + 0.1656*dist

```{r}
plot(cars)
```

```{r}
cor(cars)
```

```{r}
summary(model)
```
Looking at the model, we observe that the F-statistics p value is 1.49*10^-12 which makes it significant in the 99.99% confidence interval or almost 100% confidence interval. Hence we can reject null hypothesis and assume the model is statistically significant

Also, the p-value for t-test on intercept is 1.44*10^-12, which tells us that it is significant in the 99.99% confidence interval. hence the intercept is statistically significant

Also, the p-value for t-test on coeff of dist is 1.49*10^-12, which tells us that it is significant in the 99.99% confidence interval. hence the coefficient is statistically significant

the goodness of fit of the model here is 0.6511 based on multiple R-sq

The resultant equation is Y=8.2839 + 0.1656*dist
